# Introduction to the Tidyverse

## Data wrangling

### Load dataset

```{r}
# Load the gapminder package
library(gapminder)

# Load the dplyr package
library(dplyr)

# Look at the gapminder dataset
gapminder
```

### Filtering

The `filter` verb extracts particular observations based on a condition. pipe (`%>%`)

```{r}
# Filter the gapminder dataset for the year 1957
# version 1
gapminder %>%
    filter(year == 1957)

# version 2
gapminder[gapminder$year == 1957, ]
```

Use the `filter()` verb to set *two conditions*.

```{r}
# Filter for China in 2002
# version 1
gapminder %>%
    filter(country == "China", year == 2002)

# version 2
gapminder[gapminder$country == "China" & gapminder$year == 2002, ]
```

### Arrange

Use `arrange()` to sort observations in ascending or descending order of a particular variable.

```{r}
# Sort in ascending order of lifeExp
gapminder %>%
    arrange(lifeExp)
  
# Sort in descending order of lifeExp
gapminder %>% 
    arrange(desc(lifeExp))
```

Filtering and arranging

```{r}
# Filter for the year 1957, then arrange in descending order of population
gapminder %>%
    filter(year == 1957) %>%
    arrange(desc(pop))
```

### Mutate

Can change exist columns, or create new columns.

```{r}
# Use mutate to change lifeExp to be in months
gapminder %>%
    mutate(lifeExp = 12 * lifeExp)

# Use mutate to create a new column called lifeExpMonths
gapminder %>%
    mutate(lifeExpMonths = 12 * lifeExp)
```

Combining filter, mutate, and arrange.

```{r}
# Filter, mutate, and arrange the gapminder dataset
gapminder %>%
    filter(year == 2007) %>%
    mutate(lifeExpMonths = 12*lifeExp) %>%
    arrange(desc(lifeExpMonths))
```

## Data visualization

### Subset df variable

```{r}
# Load the ggplot2 package as well
library(ggplot2)

# Create gapminder_1952
gapminder_1952 <- gapminder %>%
    filter(year == 1952)
```

### ggplot aesthetics

There are three parts to a ggplot graph.

1.  Data: The data that we're visualizing.

2.  Aesthetic mapping: The mapping of variables in your dataset to aesthetics in your graph. (In a scatterplot, your two dimensions are the x axis and the y axis)

3.  Layer: Specifying the type of graph you're creating. You do that by adding a layer to the graph - use a `+` after the ggplot, and then `geom_point()`. The "geom" means you're adding a type of geometric object to the graph, the "point" indicates it's a scatter plot, where each observation corresponds to one point.

```{r}
# Change to put pop on the x-axis and gdpPercap on the y-axis
ggplot(gapminder_1952, aes(x = pop, y = gdpPercap)) +
    geom_point()
```

```{r}
# Create a scatter plot with pop on the x-axis and lifeExp on the y-axis
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
    geom_point()
```

#### Log scales

Putting the x-axes on a log scale

Since population is spread over several orders of magnitude, with some countries having a much higher population than others, it's a good idea to put the x-axis on a log scale.

```{r}
# Change this plot to put the x-axis on a log scale
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
    geom_point() +
    scale_x_log10()
```

Putting the x- and y- axes on a log scale

```{r}
# Scatter plot comparing pop and gdpPercap, with both axes on a log scale
ggplot(gapminder_1952, aes(x = pop, y = gdpPercap)) +
    geom_point() +
    scale_x_log10() +
    scale_y_log10()
```

#### Colors

By syntax `aes(..., color = var)`

```{r}
# Scatter plot comparing pop and lifeExp, with color representing continent
ggplot(gapminder_1952, aes(x = pop, y = lifeExp, color = continent)) +
    geom_point() +
    scale_x_log10()
```

#### Size

By syntax `aes(..., size = var)`

```{r}
# Add the size aesthetic to represent a country's gdpPercap
ggplot(gapminder_1952, aes(x = pop, y = lifeExp, color = continent,
                           size = gdpPercap)) +
    geom_point() +
    scale_x_log10()
```

### Faceting

ggplot2 lets you divide your plot into subplots to get one smaller graph for each factor of categorical variables. By syntax `facet_wrap(~ var)`.

```{r}
# Scatter plot comparing pop and lifeExp, faceted by continent
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
    geom_point() +
    scale_x_log10() +
    facet_wrap(~ continent)
```

Faceting by year

Now that you're able to use faceting, however, you can create a graph showing all the country-level data from 1952 to 2007, to understand how global statistics have changed over time.

```{r}
# Scatter plot comparing gdpPercap and lifeExp, with color representing continent
# and size representing population, faceted by year
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent,
                      size = pop)) +
    geom_point() +
    scale_x_log10() +
    facet_wrap(~ year)
```

## Grouping & summarizing

### Summarize

Turn many rows into one.(變項的summary) Fuctions you can use in `summarize()`: mean, sum, median, min, max.

Summarize multiple variables at once:

```{r}
# Filter for 1957 then summarize the median life expectancy and the maximum GDP per capita
gapminder %>% 
    filter(year == 1957) %>%
    summarize(medianLifeExp = median(lifeExp),
              maxGdpPercap = max(gdpPercap))
```

### Group by

Turn groups into one row each.

Summarizing by one variable.

```{r}
# Find median life expectancy and maximum GDP per capita in each year
gapminder %>%
    group_by(year) %>%
    summarize(medianLifeExp = median(lifeExp),
              maxGdpPercap = max(gdpPercap))
```

```{r}
# Find median life expectancy and maximum GDP per capita in each continent in 1957
gapminder %>%
    filter(year == 1957) %>%
    group_by(continent) %>%
    summarize(medianLifeExp = median(lifeExp),
              maxGdpPercap = max(gdpPercap))
```

Summarizing by more variables.

```{r}
# Find median life expectancy and maximum GDP per capita in each continent/year combination
gapminder %>% 
    group_by(year, continent) %>%
    summarize(medianLifeExp = median(lifeExp),
              maxGdpPercap = max(gdpPercap))
```

### Visualizing summarized data

Add `expand_limits(y = 0)` to make sure the plot's y-axis includes zero.

```{r}
# Create a scatter plot showing the change in medianLifeExp over time
by_year <- gapminder %>%
    group_by(year) %>%
    summarize(medianLifeExp = median(lifeExp),
              maxGdpPercap = max(gdpPercap))

ggplot(by_year, aes(x = year, y = medianLifeExp)) +
    geom_point() +
    expand_limits(y = 0)
```

```{r}
# Summarize medianGdpPercap within each continent within each year: by_year_continent
by_year_continent <- gapminder %>% 
    group_by(continent, year) %>%
    summarize(medianGdpPercap = median(gdpPercap))

# Plot the change in medianGdpPercap in each continent over time
ggplot(by_year_continent, aes(x = year, y = medianGdpPercap, 
                              color = continent)) +
    geom_point() +
    expand_limits(y = 0)
```

```{r}
# Summarize the median GDP and median life expectancy per continent in 2007
by_continent_2007 <- gapminder %>%
    filter(year == 2007) %>%
    group_by(continent) %>%
    summarize(medianLifeExp = median(lifeExp),
              medianGdpPercap = median(gdpPercap))

# Use a scatter plot to compare the median GDP and median life expectancy
ggplot(by_continent_2007, aes(x = medianGdpPercap, y = medianLifeExp,
                              color = continent)) +
    geom_point()
```

## Types of visualizations

|           | line plots                          | bar plots                                           | histograms                                                      | box plots                                                               |
|--------|--------------|---------------|------------------|-------------------|
| **usage** | useful for showing change over time | comparing statistics for each of several categories | describe the distribution of a one-dimensional numeric variable | compare the distribution of a numeric variable among several categories |
| **geom**  | `geom_line()`                       | `geom_col()`                                        | `geom_histogram(bins = int)`                                    | `geom_boxplot()`                                                        |

### Line plots

A line plot is useful for visualizing trends over time.
```{r}
# Summarize the median gdpPercap by year, then save it as by_year
by_year <- gapminder %>%
    group_by(year) %>%
    summarize(medianGdpPercap = median(gdpPercap))

# Create a line plot showing the change in medianGdpPercap over time
ggplot(by_year, aes(x = year, y = medianGdpPercap)) +
    geom_line() +
    expand_limits(y = 0)
```
```{r}
# Summarize the median gdpPercap by year & continent, save as by_year_continent
by_year_continent <- gapminder %>%
    group_by(year, continent) %>%
    summarize(medianGdpPercap = median(gdpPercap))

# Create a line plot showing the change in medianGdpPercap by continent over time
ggplot(by_year_continent, aes(x = year, y = medianGdpPercap,
                              color = continent)) +
    geom_line() +
    expand_limits(y = 0)
```

### Bar plots

A bar plot is useful for visualizing summary statistics.
```{r}
# Summarize the median gdpPercap by continent in 1952
by_continent <- gapminder %>%
    filter(year == 1952) %>%
    group_by(continent) %>%
    summarise(medianGdpPercap = median(gdpPercap))

# Create a bar plot showing medianGdp by continent
ggplot(by_continent, aes(x = continent, y = medianGdpPercap)) +
    geom_col()
```
```{r}
# Filter for observations in the Oceania continent in 1952
oceania_1952 <- gapminder %>%
    filter(year == 1952 & continent == "Oceania")

# Create a bar plot of gdpPercap by country
ggplot(oceania_1952, aes(x = country, y = gdpPercap)) +
    geom_col()
```

### Histograms

A histogram is useful for examining the distribution of a numeric variable. So there is only x-axis.

```{r}
gapminder_1952 <- gapminder %>%
    filter(year == 1952) %>%
    mutate(pop_by_mil = pop / 1000000)

# Create a histogram of population (pop_by_mil)
ggplot(gapminder_1952, aes(x = pop_by_mil)) +
    geom_histogram(bins = 50)
```

To make the histogram more informative, you can try putting the x-axis on a log scale.
```{r}
gapminder_1952 <- gapminder %>%
    filter(year == 1952)

# Create a histogram of population (pop), with x on a log scale
ggplot(gapminder_1952, aes(x = pop)) +
    geom_histogram(bins = 40) +
    scale_x_log10()
```

### Box plots

A boxplot is useful for comparing a distribution of values across several groups. 

```{r}
gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Create a boxplot comparing gdpPercap among continents
ggplot(gapminder_1952, aes(x = continent, y = gdpPercap)) +
    geom_boxplot() +
    scale_y_log10()
```















